Dirac neutrino mass generation from Majorana messenger

### 2019-08-15 draft.tex
Updates from Overleaf
jfjf

With the aim to illustrate the one-loop Dirac neutrino mass generation
we consider the particle content shown in Table \ref{tab:partcont}. as
a possible realization of the effective operator in
ec.~\eqref{eq:ld6}. Specifically, we introduce three scalar fields
$\eta, \sigma$ and $S$, where only $S$ develops a nonzero vacuum
expectation value (VEV),  two sets of three singlet fermions,
$\nu_{Ri}$ and $\nu_{R3}$, and a set a heavy Majorana fermions, $\psi_{R\alpha}$, where $i=1,2$ and
$\alpha=1,2,3$.  The model also includes an additional $\operatorname{U}(1)_X$ gauge
symmetry which is used to forbid the Dirac and Majorana neutrino mass
terms at tree level.
%Additionally, we add three right neutrinos, one scalar doublet, two scalar singlets and three Majorana fermion singlet under $\operatorname{SU}(2)_L$.
The $\operatorname{U}(1)_X$ charges for the new particles are defined by the anomaly cancellation conditions and the gauge invariance in Yukawa and scalar interactions.
%In this model the neutrinos are Dirac fermions. Particle content and their respective charges are show in Tab.~\ref{tab:partcont}.

@@

@@

A not-so-minimal model can be obtained as the one-loop realization of
the effective 5-dimensional operator for Dirac neutrino masses~\cite{Gu:2006dc}
\begin{align}
  \label{eq:ld5}
  \mathcal{L}_{5D}=\frac{1}{\Lambda} \overline{L_i} H \nu_R S\,,
\end{align}
as displayed in fig.~\ref{fig:ld5}.
In this case we can use the solutions to the anomaly cancellation conditions found for an Abelian gauge symmetry $\operatorname{U(1)_X}$ in~\cite{Campos:2017dgc}.
We use $f$ ($f$) to denote the $X$-charge of the field $f_R$ ($F_L$) for one generation of the SM. The three linear anomalies in $\operatorname{U(1)_X}$ allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $\alpha$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \sum_{\alpha} n_{\alpha}+3 (e-2l)=&0\,, &   \sum_{\alpha} n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}

We are interested here in the case of the exotic set of $X$-charges
charges: $\mp 4$,~$\mp 4$,~$\pm 5$, that when assigned to the
three light right-handed neutrinos can forbid three level Dirac and
Majorana masses~\cite{Calle:2018ovc}. This set is compatible with the
Abelian gauge groups $X=B-L,A,B,D,G$ defined in~\cite{Campos:2017dgc}.
The presence of a set of heavy Majorana mediators to realize the
dimension-5 operator at one-loop with the diagram in
fig.~\ref{fig:ld5}., would spoil the anomaly cancellation conditions.
However, we can further add an extra set of quiral fermion in a
vector-like way, such that the full set heavy fermions do not affect
the anomaly cancellation. A such assignment is displayed in
Table~\ref{tab:partcont2}.
The fields $\xi_{L \alpha}$ guarantee the anomaly cancellation and
their Majorana masses are obtained in the same way than the ones for
$\psi_{R\alpha}$.

@@

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ & $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_{B-L}$& $\operatorname{U}(1)_A$& $\operatorname{U}(1)_B$& $\operatorname{U}(1)_D$& $\operatorname{U}(1)_G$\\ \hline
$d_R $  & $\boldsymbol{1}$& $-1/2$  &  $1/3$&  $-1$&  $1$&  $0$&  $2/3$ \\
$u_R $  & $\boldsymbol{1}$& $+2/3$  &  $1/3$&  $1$&  $-1$&  $1$&  $-1/3$ \\
$Q $  & $\boldsymbol{2}$& $-1/6$  &  $1/3$&  $0$ &  $0$&  $1/2$&  $1/6$ \\\hline
$e_R $  & $\boldsymbol{1}$& $-1$  &  $-1$&  $1$ &  $-1$ &  $-2$&  $0$ \\
$L $  & $\boldsymbol{2}$& $-1/2$  &  $-1$&  $0$ &  $0$ &  $-3/2$&  $-1/2$ \\
$H $  & $\boldsymbol{2}$& $-1/2$  &  $0$&  $1$ &  $-1$ &  $1/2$&  $-1/2$ \\
$\eta$  & $\boldsymbol{2}$ & $+1/2$  & $5/2$& $-3/2$ &$3/2$ &$3$&$2$ \\
$S$ & $\boldsymbol{1}$ & $0$  & $3$& $-3$ & $3$ & $3$& $3$ \\
$\sigma$ & $\boldsymbol{1}$ & $0$ & $5/2$& $-5/2$ & $5/2$ & $5/2$& $5/2$ \\
\hline
 $\nu_{Ri}$ & $\boldsymbol{1}$ & $0$ & $-4$& $4$ & $-4$ & $-4$& $-4$\\
$\nu_{R3}$ & $\boldsymbol{1}$ & $0$ & $+5$& $-5$ & $5$ & $5$& $5$\\
$\psi_{R\alpha}$  & $\boldsymbol{1}$ & $0$ & $3/2$& $-3/2$ & $3/2$ & $3/2$& $3/2$ \\\hline
$\xi_{L\alpha}$  & $\boldsymbol{1}$ & $0$ & $3/2$ & $-3/2$ & $3/2$ & $3/2$& $3/2$\\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges. The SM fields have the usual $U(1)_{B-L}$ assignment. Now $\alpha=1,2$}
    \label{tab:partcont2}
\end{table}

>>>>

eszWith the aim to illustrate the one-loop Dirac neutrino mass generation
we consider the particle content shown in Table \ref{tab:partcont}. as
a possible realization of the effective operator in
ec.~\eqref{eq:ld6}. Specifically, we introduce three scalar fields
$\eta, \sigma$ and $S$, where only $S$ develops a nonzero vacuum
expectation value (VEV),  two sets of three singlet fermions,
$\nu_{Ri}$ and $\nu_{R3}$, and a set a heavy Majorana fermions, $\psi_{R\alpha}$, where $i=1,2$ and
$\alpha=1,2,3$.  The model also includes an additional $\operatorname{U}(1)_X$ gauge
symmetry which is used to forbid the Dirac and Majorana neutrino mass
terms at tree level.
%Additionally, we add three right neutrinos, one scalar doublet, two scalar singlets and three Majorana fermion singlet under $\operatorname{SU}(2)_L$.
The $\operatorname{U}(1)_X$ charges for the new particles are defined by the anomaly cancellation conditions and the gauge invariance in Yukawa and scalar interactions.
%In this model the neutrinos are Dirac fermions. Particle content and their respective charges are show in Tab.~\ref{tab:partcont}.

@@

For $i \neq j$ we have CLFV processes, these processes are severely restricted by the current bounds , which set limits to the Yukawas. The most restrictive experimental bounds is decay $\mu \to e \gamma$, which is several orders of magnitude greater than radiative decays $\tau \to~l\gamma$, with $l = e, \mu$. These restrictions are given by Br$(\mu \to e\gamma)$ \textless $5.7 \times 10^{-13} $~\

@@

A not-so-minimal model can be obtained as the one-loop realization of
the effective 5-dimensional operator for Dirac neutrino masses~\cite{Gu:2006dc}
\begin{align}
  \label{eq:ld5}
  \mathcal{L}_{5D}=\frac{1}{\Lambda} \overline{L_i} H \nu_R S\,,
\end{align}
as displayed in fig.~\ref{fig:ld5}.
In this case we can use the solutions to the anomaly cancellation conditions found for an Abelian gauge symmetry $\operatorname{U(1)_X}$ in~\cite{Campos:2017dgc,Okada:2018tgy}.
We use $f$ ($f$) to denote the $X$-charge of the field $f_R$ ($F_L$) for one generation of the SM. The three linear anomalies in $\operatorname{U(1)_X}$ allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $\alpha$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \sum_{\alpha} n_{\alpha}+3 (e-2l)=&0\,, &   \sum_{\alpha} n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
There are many known solutions for the case in which $e-2l=1$. The full set of standard model charges in terms of $l$ is just
\begin{align}
  u=&-1-\frac{4l}{3}\,,&d=&1+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&1+2l\,,&h=&-1-l\,.
\end{align}
The same solution but in terms of $h$ is given in ~\cite{Okada:2018tgy}.

For this kind of solutions we have that
\begin{align}
  \psi=&-\frac{\nu+1}{2}\,,&\eta=&-\frac{\nu+1}{2}-l\,,&\sigma=&\frac{1-\nu}{2}\,.
\end{align}

We are interested here in the case of the exotic set of $X$-charges
charges: $\mp 4$,~$\mp 4$,~$\pm 5$, that when assigned to the
three light right-handed neutrinos can forbid three level Dirac and
Majorana masses~\cite{Calle:2018ovc}. This set is compatible with the
Abelian gauge groups $X=B-L,B,D,G$ defined in~\cite{Campos:2017dgc}.
The presence of a set of heavy Majorana mediators to realize the
dimension-5 operator at one-loop with the diagram in
fig.~\ref{fig:ld5}., would spoil the anomaly cancellation conditions.
However, we can further add an extra set of quiral fermion in a
vector-like way, such that the full set heavy fermions do not affect
the anomaly cancellation. A such assignment is displayed in
Table~\ref{tab:partcont2}.
The fields $\xi_{L \alpha}$ guarantee the anomaly cancellation and
their Majorana masses are obtained in the same way than the ones for
$\psi_{R\alpha}$.

@@

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ & $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_{X}$& $\operatorname{U}(1)_{B-L}$& $\operatorname{U}(1)_B$& $\operatorname{U}(1)_D$& $\operatorname{U}(1)_G$\\ \hline
$L $  & $\boldsymbol{2}$& $-1/2$  &  $l$&  $-1$&    $0$ &  $-3/2$&  $-1/2$ \\    
$d_R $  & $\boldsymbol{1}$& $-1/2$  &  $1+2l/3$&  $1/3$&    $1$&  $0$&  $2/3$ \\
$u_R $  & $\boldsymbol{1}$& $+2/3$  &  $-1-4l/3$&  $1/3$&   $-1$&  $1$&  $-1/3$ \\
$Q $  & $\boldsymbol{2}$& $-1/6$  &  $-l/3$& $1/3$&    $0$&  $1/2$&  $1/6$ \\\hline
$e_R $  & $\boldsymbol{1}$& $-1$  &  $1+2l$&  $-1$&    $1$ &  $-2$&  $0$ \\
$H $  & $\boldsymbol{2}$& $-1/2$  &  $-1-l$&  $0$&    $-1$ &  $1/2$&  $-1/2$ \\
$\eta$  & $\boldsymbol{2}$ & $+1/2$  & $3/2-l$& $5/2$& $3/2$ &$3$&$2$ \\
$S$ & $\boldsymbol{1}$ & $0$  & $3$& $3$&  $3$ & $3$& $3$ \\
$\sigma$ & $\boldsymbol{1}$ & $0$ & $5/2$& $5/2$&  $5/2$ & $5/2$& $5/2$ \\
\hline
 $\nu_{Ri}$ & $\boldsymbol{1}$ & $0$ & $-4$& $-4$&  $-4$ & $-4$& $-4$\\
$\nu_{R3}$ & $\boldsymbol{1}$ & $0$ & $+5$& $+5$&  $5$ & $5$& $5$\\
$\psi_{R\alpha}$  & $\boldsymbol{1}$ & $0$& $3/2$ & $3/2$&  $3/2$ & $3/2$& $3/2$ \\\hline
$\xi_{L\alpha}$  & $\boldsymbol{1}$ & $0$ & $3/2$& $3/2$ &  $3/2$ & $3/2$& $3/2$\\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges. The SM fields have the usual $U(1)_{B-L}$ assignment. Now $\alpha=1,2$}
    \label{tab:partcont2}
\end{table}
